Diversity dynamics in Nymphalidae butterflies: Effect of phylogenetic uncertainty on diversification rate shift estimates

### 2014-10-29 MS_revised.tex
figures

\documentclass[]{article}
\usepackage{setspace}
\usepackage[]{lineno}
\def\linenumberfont{\normalfont\small\sffamily}
\doublespacing
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{tabu}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage[none]{hyphenat}
\usepackage{fixltx2e} % provides \textsubscript
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
\usepackage[margin=1in,a4paper]{geometry}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={},
            colorlinks=true,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{0}

\linenumbers

\author{}
\date{}

@@

\textbf{Title: Diversity dynamics in Nymphalidae butterflies: Effect of
phylogenetic uncertainty on diversification rate shift estimates}

\subsubsection{Authors: Carlos Peña\textsuperscript{1} and Marianne
Espeland\textsuperscript{2}}\label{authors-carlos-peuxf1a1-and-marianne-espeland2}

\subsubsection{Running title: Diversity dynamics in Nymphalidae
butterflies}\label{running-title-diversity-dynamics-in-nymphalidae-butterflies}

\emph{\textsuperscript{1} Laboratory of Genetics, Department of Biology,
University of Turku, Turku, Finland. Email:
\href{mailto:mycalesis@gmail.com}{\nolinkurl{mycalesis@gmail.com}}}

\emph{\textsuperscript{2} Museum of Comparative Zoology and Department
of Organismic and Evolutionary Biology, Harvard University, Cambridge,
USA. Email:
\href{mailto:marianne.espeland@gmail.com}{\nolinkurl{marianne.espeland@gmail.com}}.
Current address: McGuire Center for Lepidoptera and Biodiversity,
Florida Museum of Natural History, University of Florida, Gainesville,
FL, US}.

\textbf{Corresponding author:}
\href{mailto:mycalesis@gmail.com}{\nolinkurl{mycalesis@gmail.com}}
Telephone: +358 417063065 Fax: +358 2 333 6680

\textbf{Comments:} 5814 words, 7 figures, 2 tables and 25 supplementary
material files.

\textbf{Additional Supplementary materials:}
\url{http://dx.doi.org/10.6084/m9.figshare.639208}

\subsection{Abstract}\label{abstract}

@@

\subsection{Introduction}\label{introduction}

@@

\subsection{Methods}\label{methods}

\subsubsection{Data}\label{data}

@@

\subsubsection{Analyses of
Diversification}\label{analyses-of-diversification}

@@

\subsubsection{Detecting diversification shifts on phylogenetic
trees}\label{detecting-diversification-shifts-on-phylogenetic-trees}

@@

\subsubsection{Estimation of trait-dependent speciation
rates}\label{estimation-of-trait-dependent-speciation-rates}

@@

\subsection{Results}\label{results}

\subsubsection{Detecting diversification shifts on the maximum clade
credibility
tree}\label{detecting-diversification-shifts-on-the-maximum-clade-credibility-tree}

@@

\subsubsection{Phylogenetic uncertainty in the MultiMEDUSA
approach}\label{phylogenetic-uncertainty-in-the-multimedusa-approach}

@@

\subsubsection{Estimation of trait-dependent speciation
rates}\label{estimation-of-trait-dependent-speciation-rates-1}

@@

\subsection{Discussion}\label{discussion}

\subsubsection{Effects of phylogenetic uncertainty on the performance of
MEDUSA}\label{effects-of-phylogenetic-uncertainty-on-the-performance-of-medusa}

@@

\subsubsection{Hostplant use and diversification in
Nymphalidae}\label{hostplant-use-and-diversification-in-nymphalidae}

\paragraph{Ithomiini}\label{ithomiini}

@@

\paragraph{Danaini}\label{danaini}

Our MultiMEDUSA approach showed a significant slowdown in net
diversification rate in the subtribe Danaina of the Danini. Both Danaina
and the sister clade Euploeina feed mainly on Apocynaceae and thus a
hostplant shift should not be responsible for the observed slowdown of
diversification in the Danaina. As expected, our BiSSE analysis of
Apocynaceae feeders shows that there is no effect of feeding on this
plant family on the net diversification rates of Nymphalidae lineages.
Many of the Danaina are large, strong fliers, highly migratory and
involved in mimicry rings. Among them is for example the monarch
(\emph{Danaus plexippus}), probably the most well known of all migratory
butterflies. The causes for a lower net diversification rate in the
Danaina remains to be investigated, but their great dispersal power
might be involved in preventing allopatric speciation. It has been found
in highly vagile species in the nymphalid genus \emph{Vanessa} that
dispersal has homogenized populations due to gene flow, as old and
vagile species seem to be genetically homogeneous while younger
widespread species show higher genetic differentiation in their
populations {[}50{]}.

\paragraph{Satyrini}\label{satyrini}

@@

\subsection{Conclusions}\label{conclusions}

We found that even though MEDUSA estimated several diversification
shifts in the maximum clade credibility tree of Nymphalidae, only a few
of these splits were found in more than 90\% of the trees from the
posterior distribution. In the literature, it is common practice that
conclusions are based on the splits estimated on the maximum clade
credibility tree. However, by using a MultiMEDUSA approach, we found
that for this Nymphalidae dataset some of these splits might be greatly
affected by phylogenetic uncertainty. Moreover, some of these splits can
be recovered either as increases or decreases in net diversification
rate depending on the tree from the posterior distribution that was used
for analysis. This means that contradictory conclusions would be made if
only the maximum clade credibility tree was used for analysis. We
recommend that all datasets should be analyzed using both approaches,
MEDUSA and MultiMEDUSA, in order to test whether the results are robust
when phylogenetic uncertainty is taken into account.

@@

\subsection{Acknowledgments}\label{acknowledgments}

We are thankful to Mark Cornwall for help with the script to extend
MEDUSA to include phylogenetic uncertainty, Niklas Wahlberg for
commenting on the manuscript and giving us the posterior distribution of
trees, Luke Harmon for commenting on the manuscript and anonymous
reviewers for their comments, which greatly improved the manuscript,
Jessica Slove Davidson and Niklas Janz for access to their hostplant
data. The study was supported by a Kone Foundation grant (awarded to
Niklas Wahlberg), Finland (C. Peña) and the Research Council of Norway
(grant no. 204308 to M. Espeland). We acknowledge CSC--IT Center for
Science Ltd. (Finland) for the allocation of computational resources.

\subsection{Figure legends}\label{figure-legends}

\texttt{Figure 1.} Results of the MEDUSA analysis run on the maximum
clade credibility tree from {[}13{]}. Rate shifts were estimated for the
following nodes (besides the background rate): 2) Limenitidinae +
Heliconiinae, 3) \emph{Ypthima}, 4) \emph{Charaxes}, 5) Ithomiini in
part, 6) Satyrini, 7) Coenonymphina 8) Phyciodina in part, 9) Danaini in
part, 10) \emph{Coenonympha}, 11) \emph{Caeruleuptychia} +
\emph{Magneuptychia}, 12) \emph{Callicore} + \emph{Diaethria}, 13)
Satyrina, 14) Mycalesina, 15) \emph{Pedaliodes}, 16) \emph{Dryas} +
\emph{Dryadula}, 17) \emph{Taenaris}, 18) Pseudergolinae. Circles on
nodes indicate the diversification shift number as found by MEDUSA.
Numbers next to circles indicat the posterior probability values for
such nodes.

\texttt{Figure 2.} Diversification rates for taxa estimated by MEDUSA on
the samples of 1000 random trees.

\texttt{Figure 3.} Results of MultiMEDUSA analysis showing the
probability of specific nodes being characterized by significant shifts
in diversification rate.

\texttt{Figure 4.} BiSSE analysis of diversification of nymphalids due
to feeding on Solanaceae hostplants. Speciation and net diversification
rates are significantly higher in Solanaceae feeders (speciation rate =
\(\lambda1\), net diversification rate = \(r1\)).

\texttt{Table 3.} Likelihood ratio test between the model of increased
diversification of nymphalids feeding on Solanaceae against a model
forcing equal speciation rates (no effect on diversification).

\section{Supp. mat.}\label{supp.-mat.}

\texttt{Supporting\_Information\_S01.nex}: MCC Nymphalidae tree from
{[}13{]}.

\texttt{Supporting\_Information\_S02.R}: R script to run MEDUSA on the
MCC Nymphalidae tree from {[}13{]}. This script removes the outgroup
taxa and loads the richness data for the tree terminals.

@@

\texttt{Figure S1.} Boxplot of speciation (\(\lambda\)) and extinction
(\(\mu\)) values for the 95\% credibility intervals of values estimated
by BiSSE analysis of diversification due to feeding on Solanaceae
plants.

\texttt{Figure S2.} BiSSE analysis of diversification of nymphalids due
to feeding on Solanaceae hostplants assuming \emph{Vanessa} and
\emph{Hypanartia} as non-Solanaceae feeders. The same pattern is
recovered, speciation and net diversification rates are significantly
higher for Solanaceae feeders (\(\lambda1\), \(r1\)).

@@

\section*{References}\label{references}
\addcontentsline{toc}{section}{References}

1. Mitter C, Farrell B, Wiegmann B (1988) The phylogenetic study of
adaptive zones: Has phytophagy promoted insect diversification? Am Nat
132: 107--128.
doi:\href{http://dx.doi.org/10.1086/284840}{10.1086/284840}.

@@

14. Peña C, Wahlberg N (2008) Prehistorical climate change increased
diversification of a group of butterflies. Biol Lett 4: 274--278.
doi:\href{http://dx.doi.org/10.1098/rsbl.2008.0062}{10.1098/rsbl.2008.0062}.

@@

\section*{Tables}
\begin{table}[!h]
\caption{\bf{Significant net diversification rate shifts found in the MEDUSA analysis of the Nymphalid maximum clade credibility tree.}}
\begin{tabular}{lccrrcl}
Shift N$^\circ$ & Split.Node & Model & r          & LnLik.part & AICc     & Taxa                                                       \\
1               & 399        & yule  & 0.092459   & -1055.957  &          & Nymphalidae (root)                                         \\
2               & 691        & bd    & 0.054129   & -406.3703  &          & Limenitidinae + Heliconiinae                               \\
3               & 299        & yule  & 0.311199   & -6.3058    &          & \emph{Ypthima}                                             \\
4               & 224        & yule  & 0.290989   & -6.2601    &          & \emph{Charaxes}                                            \\
5               & 750        & yule  & 0.186913   & -147.4146  &          & Oleriina + Ithomiina + Napeogenina + Dircennina + Godyrina \\
6               & 405        & yule  & 0.116252   & -555.0276  &          & Satyrini                                                   \\
7               & 495        & yule  & 0.064656   & -124.9143  &          & Coenonymphina                                              \\
8               & 609        & yule  & 0.240562   & -35.0908   &          & Phyciodina in part                                         \\
9               & 787        & bd    & 0.042332   & -43.7819   &          & Danaini in part                                            \\
10              & 231        & yule  & 0.209416   & -4.7955    &          & \emph{Coenonympha}                                         \\
11              & 478        & yule  & 0.311684   & -9.2218    &          & \emph{Caeruleuptychia} + \emph{Magneuptychia}              \\
12              & 659        & yule  & 0.219253   & -11.2686   &          & \emph{Callicore} + \emph{Diaethria}                        \\
13              & 444        & yule  & 0.220615   & -43.7812   &          & Satyrina                                                   \\
14              & 524        & yule  & 0.190754   & -26.0651   &          & Mycalesina                                                 \\
15              & 355        & yule  & 0.234041   & -6.2013    &          & \emph{Pedaliodes}                                          \\
16              & 714        & yule  & 0          & 0          &          & \emph{Dryas} + \emph{Dryadula}                             \\
17              & 377        & yule  & 0.311671   & -4.1986    &          & \emph{Taenaris}                                            \\
18              & 688        & yule  & 0.024724   & -17.5256   &          & Pseudergolinae                                             \\
19              & 583        & yule  & 0          & 0          & 5090.492 & \emph{Anaeomorpha} + \emph{Hypna}                                       
\end{tabular}
\begin{flushleft}Split.Node=node number, Model=preferred diversification model by MEDUSA, r=net diversification rate, LnLik.part=log likelihood value.
\end{flushleft}
\end{table}

\begin{table}[!h]
    \caption{\bf{Differences in rates estimated by MEDUSA on the MCC
tree from Wahlberg and the MultiMEDUSA approach on 1000 random trees
from the posterior distribution.}}
\begin{tabular}{lcrr}
Split.Node & rate by MEDUSA & Median rate by MultiMEDUSA & probability of being recovered \\
1          & 0.090          & not found                  & 0.000                          \\
2          & 0.054          & -0.030                     & 0.864                          \\
3          & 0.311          & 0.219                      & 0.961                          \\
4          & 0.290          & 0.212                      & 0.996                          \\
5          & 0.187          & 0.101                      & 0.927                          \\
6          & 0.116          & 0.039                      & 0.131                          \\
7          & 0.064          & -0.052                     & 0.195                          \\
8          & 0.240          & 0.166                      & 0.319                          \\
9          & 0.042          & -0.049                     & 0.897                          \\
10         & 0.209          & 0.149                      & 0.619                          \\
11         & 0.311          & 0.208                      & 0.831                          \\
12         & 0.219          & 0.135                      & 0.911                          \\
13         & 0.220          & 0.117                      & 0.276                          \\
14         & 0.190          & 0.115                      & 0.379                          \\
15         & 0.234          & 0.127                      & 0.825                          \\
16         & 0.000          & -0.005                     & 0.659                          \\
17         & 0.311          & 0.248                      & 0.785                          \\
18         & 0.024          & -0.064                     & 0.024                          \\
19         & 0.000          & -0.087                     & 0.381                         
\end{tabular}
\end{table}
% 
% See introductory notes if you wish to include sideways tables.
%
% NOTE: Please look over our table guidelines at http://www.plosone.org/static/figureGuidelines#tables to make sure that your tables meet our requirements. Certain types of spacing, cell merging, and other formatting tricks may have unintended results and will be returned for revision.
%
%\begin{table}[!ht]
%\caption{
%\bf{Table title}}
%\begin{tabular}{|c|c|c|}
%table information
%\end{tabular}
%\begin{flushleft}Table caption
%\end{flushleft}
%\label{tab:label}
% \end{table}

>>>>

\documentclass[10pt]{article}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{cite}
\usepackage{hyperref}

\usepackage{lineno}
\usepackage{microtype}
\DisableLigatures[f]{encoding = *, family = * }

% Text layout
\topmargin 0.0cm
\oddsidemargin 0.5cm
\evensidemargin 0.5cm
\textwidth 16cm 
\textheight 21cm

% Bold the 'Figure #' in the caption and separate it with a period
% Captions will be left justified
\usepackage[labelfont=bf,labelsep=period,justification=raggedright]{caption}

% Use the PLoS provided BiBTeX style
\bibliographystyle{plos2009}

% Remove brackets from numbering in List of References
\makeatletter
\renewcommand{\@biblabel}[1]{\quad#1.}
\makeatother

% Leave date blank
\date{}

\pagestyle{myheadings}

%% Include all macros below. Please limit the use of macros.

%% END MACROS SECTION

@@

\begin{flushleft}
{\Large
    \title{Diversity dynamics in Nymphalidae butterflies: Effect of phylogenetic uncertainty on diversification rate shift estimates}
}
\maketitle
Carlos Pe\~na$^{1,\ast}$,
Marianne Espeland$^{2}$
\\
\bf{1} Carlos Pe\~na Laboratory of Genetics, Department of Biology, University of Turku, Turku, Finland. Email: \href{mailto:mycalesis@gmail.com}{\nolinkurl{mycalesis@gmail.com}}
\\
\bf{2} Marianne Espeland Museum of Comparative Zoology and Department
of Organismic and Evolutionary Biology, Harvard University, Cambridge,
USA. Email:
\href{mailto:marianne.espeland@gmail.com}{\nolinkurl{marianne.espeland@gmail.com}}.
Current address: McGuire Center for Lepidoptera and Biodiversity,
Florida Museum of Natural History, University of Florida, Gainesville,
FL, US.
\\
$\ast$ \textbf{Corresponding author:}
\href{mailto:mycalesis@gmail.com}{\nolinkurl{mycalesis@gmail.com}}
Telephone: +358 417063065 Fax: +358 2 333 6680

\end{flushleft}

\subsubsection*{Running title: Diversity dynamics in Nymphalidae
butterflies}

\subsection*{Abstract}

@@

\subsection*{Introduction}

@@

\subsection*{Materialas and Methods}

\subsubsection*{Data}

@@

\subsubsection*{Analyses of
Diversification}

@@

\subsubsection*{Detecting diversification shifts on phylogenetic
trees}

@@

\subsubsection*{Estimation of trait-dependent speciation
rates}

@@

\subsection*{Results}

\subsubsection*{Detecting diversification shifts on the maximum clade
credibility
tree}

@@

\subsubsection*{Phylogenetic uncertainty in the MultiMEDUSA
approach}

@@

\subsubsection*{Estimation of trait-dependent speciation
rates}

@@

\subsection*{Discussion}

\subsubsection*{Effects of phylogenetic uncertainty on the performance of
MEDUSA}

@@

\subsubsection*{Hostplant use and diversification in
Nymphalidae}

\paragraph*{Ithomiini}

@@

\paragraph*{Danaini}
Our MultiMEDUSA approach showed a significant slowdown in net
diversification rate in the subtribe Danaina of the Danini. Both Danaina
and the sister clade Euploeina feed mainly on Apocynaceae and thus a
hostplant shift should not be responsible for the observed slowdown of
diversification in the Danaina. As expected, our BiSSE analysis of
Apocynaceae feeders shows that there is no effect of feeding on this
plant family on the net diversification rates of Nymphalidae lineages.
Many of the Danaina are large, strong fliers, highly migratory and
involved in mimicry rings. Among them is for example the monarch
(\emph{Danaus plexippus}), probably the most well known of all migratory
butterflies. The causes for a lower net diversification rate in the
Danaina remains to be investigated, but their great dispersal power
might be involved in preventing allopatric speciation. It has been found
in highly vagile species in the nymphalid genus \emph{Vanessa} that
dispersal has homogenized populations due to gene flow, as old and
vagile species seem to be genetically homogeneous while younger
widespread species show higher genetic differentiation in their
populations {[}50{]}.

\paragraph*{Satyrini}

@@

\subsection*{Conclusions}
We found that even though MEDUSA estimated several diversification
shifts in the maximum clade credibility tree of Nymphalidae, only a few
of these splits were found in more than 90\% of the trees from the
posterior distribution. In the literature, it is common practice that
conclusions are based on the splits estimated on the maximum clade
credibility tree. However, by using a MultiMEDUSA approach, we found
that for this Nymphalidae dataset some of these splits might be greatly
affected by phylogenetic uncertainty. Moreover, some of these splits can
be recovered either as increases or decreases in net diversification
rate depending on the tree from the posterior distribution that was used
for analysis. This means that contradictory conclusions would be made if
only the maximum clade credibility tree was used for analysis. We
recommend that all datasets should be analyzed using both approaches,
MEDUSA and MultiMEDUSA, in order to test whether the results are robust
when phylogenetic uncertainty is taken into account.

@@

\subsection*{Acknowledgments}
We are thankful to Mark Cornwall for help with the script to extend
MEDUSA to include phylogenetic uncertainty, Niklas Wahlberg for
commenting on the manuscript and giving us the posterior distribution of
trees, Luke Harmon for commenting on the manuscript and anonymous
reviewers for their comments, which greatly improved the manuscript,
Jessica Slove Davidson and Niklas Janz for access to their hostplant
data. The study was supported by a Kone Foundation grant (awarded to
Niklas Wahlberg), Finland (C. Pe\~na) and the Research Council of Norway
(grant no. 204308 to M. Espeland). We acknowledge CSC--IT Center for
Science Ltd. (Finland) for the allocation of computational resources.

\section*{Figure legends}
\begin{description}
 \item {\bf Figure 1. Results of the MEDUSA analysis run on the maximum
clade credibility tree from Wahlberg et al. (2009)} Rate shifts were estimated for the
following nodes (besides the background rate): 2) Limenitidinae +
Heliconiinae, 3) \emph{Ypthima}, 4) \emph{Charaxes}, 5) Ithomiini in
part, 6) Satyrini, 7) Coenonymphina 8) Phyciodina in part, 9) Danaini in
part, 10) \emph{Coenonympha}, 11) \emph{Caeruleuptychia} +
\emph{Magneuptychia}, 12) \emph{Callicore} + \emph{Diaethria}, 13)
Satyrina, 14) Mycalesina, 15) \emph{Pedaliodes}, 16) \emph{Dryas} +
\emph{Dryadula}, 17) \emph{Taenaris}, 18) Pseudergolinae. Circles on
nodes indicate the diversification shift number as found by MEDUSA.
Numbers next to circles indicat the posterior probability values for
such nodes.

\item {\bf Figure 2. Diversification rates for taxa estimated by MEDUSA on
the samples of 1000 random trees}

\item {\bf Figure 3. Results of MultiMEDUSA analysis showing the
probability of specific nodes being characterized by significant shifts
in diversification rate}

\item {\bf Figure 4. BiSSE analysis of diversification of nymphalids due
to feeding on Solanaceae hostplants} Speciation and net diversification
rates are significantly higher in Solanaceae feeders (speciation rate =
$\lambda1$, net diversification rate = $r1$.
\end{description}

\section*{Supporting Information Legends}
\begin{description}
 \item {\bf Supporting\_Information\_S01.nex.} MCC Nymphalidae tree from
{[}13{]}
\item {\bf Supporting\_Information\_S02.R} R script to run MEDUSA on the
MCC Nymphalidae tree from {[}13{]}. This script removes the outgroup
taxa and loads the richness data for the tree terminals.
\end{description}

@@

\texttt{Figure S1.} Boxplot of speciation ($\lambda$) and extinction
(\(\mu\)) values for the 95\% credibility intervals of values estimated
by BiSSE analysis of diversification due to feeding on Solanaceae
plants.

\texttt{Figure S2.} BiSSE analysis of diversification of nymphalids due
to feeding on Solanaceae hostplants assuming \emph{Vanessa} and
\emph{Hypanartia} as non-Solanaceae feeders. The same pattern is
recovered, speciation and net diversification rates are significantly
higher for Solanaceae feeders ($\lambda1$), r1).

@@

\section*{References}
1. Mitter C, Farrell B, Wiegmann B (1988) The phylogenetic study of
adaptive zones: Has phytophagy promoted insect diversification? Am Nat
132: 107--128.
doi:\href{http://dx.doi.org/10.1086/284840}{10.1086/284840}.

@@

14. Pe\~na C, Wahlberg N (2008) Prehistorical climate change increased
diversification of a group of butterflies. Biol Lett 4: 274--278.
doi:\href{http://dx.doi.org/10.1098/rsbl.2008.0062}{10.1098/rsbl.2008.0062}.

@@

% Table 1
\section*{Tables}
\begin{table}[!h]
\caption{\bf{Significant net diversification rate shifts found in the MEDUSA analysis of the Nymphalid maximum clade credibility tree.}}
\begin{tabular}{lccrrcl}
Shift N$^\circ$ & Split.Node & Model & r          & LnLik.part & AICc     & Taxa                                                       \\
1               & 399        & yule  & 0.092459   & -1055.957  &          & Nymphalidae (root)                                         \\
2               & 691        & bd    & 0.054129   & -406.3703  &          & Limenitidinae + Heliconiinae                               \\
3               & 299        & yule  & 0.311199   & -6.3058    &          & \emph{Ypthima}                                             \\
4               & 224        & yule  & 0.290989   & -6.2601    &          & \emph{Charaxes}                                            \\
5               & 750        & yule  & 0.186913   & -147.4146  &          & Oleriina + Ithomiina + Napeogenina + Dircennina + Godyrina \\
6               & 405        & yule  & 0.116252   & -555.0276  &          & Satyrini                                                   \\
7               & 495        & yule  & 0.064656   & -124.9143  &          & Coenonymphina                                              \\
8               & 609        & yule  & 0.240562   & -35.0908   &          & Phyciodina in part                                         \\
9               & 787        & bd    & 0.042332   & -43.7819   &          & Danaini in part                                            \\
10              & 231        & yule  & 0.209416   & -4.7955    &          & \emph{Coenonympha}                                         \\
11              & 478        & yule  & 0.311684   & -9.2218    &          & \emph{Caeruleuptychia} + \emph{Magneuptychia}              \\
12              & 659        & yule  & 0.219253   & -11.2686   &          & \emph{Callicore} + \emph{Diaethria}                        \\
13              & 444        & yule  & 0.220615   & -43.7812   &          & Satyrina                                                   \\
14              & 524        & yule  & 0.190754   & -26.0651   &          & Mycalesina                                                 \\
15              & 355        & yule  & 0.234041   & -6.2013    &          & \emph{Pedaliodes}                                          \\
16              & 714        & yule  & 0          & 0          &          & \emph{Dryas} + \emph{Dryadula}                             \\
17              & 377        & yule  & 0.311671   & -4.1986    &          & \emph{Taenaris}                                            \\
18              & 688        & yule  & 0.024724   & -17.5256   &          & Pseudergolinae                                             \\
19              & 583        & yule  & 0          & 0          & 5090.492 & \emph{Anaeomorpha} + \emph{Hypna}                                       
\end{tabular}
\begin{flushleft}Split.Node=node number, Model=preferred diversification model by MEDUSA, r=net diversification rate, LnLik.part=log likelihood value.
\end{flushleft}
\end{table}

% Table 2
\begin{table}[!h]
    \caption{\bf{Differences in rates estimated by MEDUSA on the MCC
tree from Wahlberg and the MultiMEDUSA approach on 1000 random trees
from the posterior distribution.}}
\begin{tabular}{lcrr}
Split.Node & rate by MEDUSA & Median rate by MultiMEDUSA & probability of being recovered \\
1          & 0.090          & not found                  & 0.000                          \\
2          & 0.054          & -0.030                     & 0.864                          \\
3          & 0.311          & 0.219                      & 0.961                          \\
4          & 0.290          & 0.212                      & 0.996                          \\
5          & 0.187          & 0.101                      & 0.927                          \\
6          & 0.116          & 0.039                      & 0.131                          \\
7          & 0.064          & -0.052                     & 0.195                          \\
8          & 0.240          & 0.166                      & 0.319                          \\
9          & 0.042          & -0.049                     & 0.897                          \\
10         & 0.209          & 0.149                      & 0.619                          \\
11         & 0.311          & 0.208                      & 0.831                          \\
12         & 0.219          & 0.135                      & 0.911                          \\
13         & 0.220          & 0.117                      & 0.276                          \\
14         & 0.190          & 0.115                      & 0.379                          \\
15         & 0.234          & 0.127                      & 0.825                          \\
16         & 0.000          & -0.005                     & 0.659                          \\
17         & 0.311          & 0.248                      & 0.785                          \\
18         & 0.024          & -0.064                     & 0.024                          \\
19         & 0.000          & -0.087                     & 0.381                         
\end{tabular}
\end{table}

% Table 3
\begin{table}[!h]
    \caption{\bf{Likelihood ratio test between the model of increased
diversification of nymphalids feeding on Solanaceae against a model
forcing equal speciation rates (no effect on diversification).}}
\begin{tabular}{lcrrlr}
Df           & lnLik & AIC     & ChiSq  & \multicolumn{1}{r}{p} &         \\
full         & 6     & -1613.3 & 3238.5 &                       &         \\
equal.lambda & 5     & -1619.4 & 3248.9 & 12.3                  & 0.00045
\end{tabular}
\end{table}
